---
title: "NoteDay3"
author: "Diamond"
date: "2020/7/8"
output: 
  html_document: default
  pdf_document:
    latex_engine: xelatex
---


# chapter 3 Visuation:ggplot2

install first package, ggplot2, by installing tidyverse


```{r setup, include=FALSE}


#install.packages("tidyverse") #Just need to do once
library(tidyverse) #when to need to use 
library(MASS)


```
Test work place simply.
```{r}

getwd()
```


## 1.Load data

```{r}

#National Parks in California
ca <- read_csv("data/ca.csv")
#Acadia National Park
acadia <- read_csv("data/acadia.csv")
#Southeast US National Parks
se <- read_csv("data/se.csv")
#2016 Visitation for all Pacific West National Parks
visit_16 <- read_csv("data/visit_16.csv")
#All Nationally designated sites in Massachusetts
mass <- read_csv("data/mass.csv")

```

## 2.A Grammar of Graphics!

ggplot(data = \<DATA\>) +
\<GEOM_FUNCTION\>(mapping = aes(\<MAPPINGS\>),stat = \<STAT\>,position = \<POSITION\>) +
\<COORDINATE_FUNCTION\> +
\<FACET_FUNCTION\>

You can uniquely describe any plot as a combination of these 7 parameters.


```{r}
head(ca)
#view(ca) other worksheet will come out
```

Among the variables in ca are:

1. region, US region where park is located.

2. visitors, the annual visitation for each year

To build a ggplot, we need to:

*use the ggplot() function and bind the plot to a specific data frame using the data argument*

```{r}
ggplot(data=ca)+
  geom_point(aes(x = year, y = visitors))
```
Notation '+' must be the last  in the line (middle also OK)

