---
title: "Homework1"
author: "Zhang YunMengGe_3170105497"
date: "2020/7/7"
output: 
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

```{r setup, message = F, include=FALSE}
#Call all packages
library(tidyverse)
library(tinytex)

```

# Question 1

The Iowa data set `iowa.csv` is a toy example that summarises the yield of wheat (bushels per acre) for the state of Iowa between 1930-1962. In addition to yield, year, rainfall and temperature were recorded as the main predictors of yield.

### a.First, we need to load the data set into R using the command `read.csv()`. Use the help function to learn what arguments this function takes. Once you have the necessary input, load the data set into R and make it a data frame called `iowa.df`.

```{r}
#Use the help function
?read.csv()

#load the data set
iowadata<-read.csv("data/Iowa.csv",header = TRUE, sep = ";")
iowa.df<-as.data.frame(iowadata)
head(iowa.df)
```

### b. How many rows and columns does `iowa.df` have? 

```{r}
#get the number of row and col separately
nrow(iowa.df)
ncol(iowa.df)

#also wo can get them directly
dim(iowa.df)
```
###  c. What are the names of the columns of `iowa.df`?

```{r}
colnames(iowa.df)
```
### d. What is the value of row 5, column 7 of `iowa.df`?
```{r}
iowa.df[5,7]
```
### e. Display the second row of `iowa.df` in its entirety.

```{r}
iowa.df[2,]
```

# Question 2: Syntax and class-typing.

###  a. For each of the following commands, either explain why they should be errors, or explain the non-erroneous result.
```
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
sum(vector1)

#[1] "7"
#[1] "12" "32" "5"  "7" 
#Error in sum(vector1) : 'type'(character)参数不对
```

The type of the elements is 'character'.

The way to find the max and to sort is according to ASCII code. Of course two characters can not be summed up.


### b. For the next series of commands, either explain their results, or why they should produce errors.
```
vector2 <- c("5",7,12)
vector2[2] + vector2[3]
#Error in vector2[2] + vector2[3] : 二进列运算符中有非数值参数
```
The type of elements in the vector should be consistent, otherwise 7 and 12 will be automatically converted into the same character vector as "5", which cannot be summed.

```{r}
dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]
```
Different types of variables can be stored in the data frame. The second and third elements in the first row are numeric vectors, which can be added together.


```
list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]
list4[2]+list4[4]

#[1] 168
#Error in list4[2] + list4[4] : 二进列运算符中有非数值参数
```
The elements of the first formula are in numerical form and therefore can be added; the elements of the second formula are in the form of a linked list and cannot be added.


# Question 3: Working with functions and operators.

### a. The colon operator will create a sequence of integers in order. It is a special case of the function `seq()` which you saw earlier in this assignment. Using the help command `?seq` to learn about the function, design an expression that will give you the sequence of numbers from 1 to 10000 in increments of 372. Design another that will give you a sequence between 1 and 10000 that is exactly 50 numbers in length.

```{r}
?seq()
seq(1,10000,by=372)

seq(1,10000,length.out = 50)
```
    
### b. The function `rep()` repeats a vector some number of times. Explain the difference between `rep(1:3, times=3) and rep(1:3, each=3).

```{r}
rep(1:3, times=3)
rep(1:3, each=3)
```
The first three-digit number "1,2,3" is generated and repeated three times as a whole.

The second produces three digits "1,2,3", where each digit is repeated three times.

# MB.Ch1.2.

The orings data frame gives data on the damage that had occurred in US space shuttle launches
prior to the disastrous Challenger launch of 28 January 1986. The observations in rows 1, 2, 4, 11, 13, and 18
were included in the pre-launch charts used in deciding whether to proceed with the launch, while remaining
rows were omitted.
Create a new data frame by extracting these rows from orings, and plot total incidents against temperature
for this new data frame. Obtain a similar plot for the full data set.

```{r}
#install.packages("DAAG")
library(DAAG)
```

```{r}
oringsdata<-data.frame(orings)
head(oringsdata)
```
```{r}
neworings<-oringsdata[c(1,2,4,11,13,18),]
head(neworings)
```

```{r}
ggplot(data = neworings) +
geom_point(aes(x = Temperature, y = Total)) +
labs(x = "Temperature",
y = "Tatal",
title = "ORINGS") 
```

# MB.Ch1.4. For the data frame ais (DAAG package)

### (a) Use the function str() to get information on each of the columns. Determine whether any of the columns hold missing values.

```{r}
library(DAAG)

str(ais)
```
```{r}
#is.na(ais)
which(is.na(ais)) 
```

### (b) Make a table that shows the numbers of males and females for each different sport. In which sports is there a large imbalance (e.g., by a factor of more than 2:1) in the numbers of the two sexes?

```{r}
ais<-data.frame(ais)
selectais=table(ais$sex,ais$sport)
selectais
```

```{r}
ais2<-t(selectais)
proportion_sport<-ais2[,1]/ais2[,2]
proportion_sport
ifimbanlance<-((ais2[,1]/ais2[,2]>2)|(ais2[,1]/ais2[,2]<0.5))
ifimbanlance
```
From the table `proportion_sport` above, we can find out the porportion of female to male.

The table `ifimbanlance` tells that if the proportion is imbalance. If yes, it turns out to be `True`, if no it turns out to be `False`.

# MB.Ch1.6.

Create a data frame called Manitoba.lakes that contains the lake’s elevation (in meters above sea
level) and area (in square kilometers) as listed below. Assign the names of the lakes using the row.names()
function. elevation area Winnipeg 217 24387 Winnipegosis 254 5374 Manitoba 248 4624 SouthernIndian 254
2247 Cedar 253 1353 Island 227 1223 Gods 178 1151 Cross 207 755 Playgreen 217 657

### (a) Use the following code to plot log2(area) versus elevation, adding labeling information (there is an extreme value of area that makes a logarithmic scale pretty much essential):

```{r}
Manitoba.lakes<-data.frame(elevation=c(217,254,248,254,253,227,178,207,217),area=c(24387,5374,4624,2247,1353,1223,1151,755,657))
row.names(Manitoba.lakes)<-c("Winnipeg","Winnipegosis","Manitoba","SouthernIndian","Cedar","Island","Gods","Cross","Playgreen")
Manitoba.lakes
```

```{r}
attach(Manitoba.lakes)
plot(log2(area) ~ elevation, pch=16, xlim=c(170,280))
# NB: Doubling the area increases log2(area) by 1.0
text(log2(area) ~ elevation, labels=row.names(Manitoba.lakes), pos=4)
text(log2(area) ~ elevation, labels=area, pos=2)
title("Manitoba’s Largest Lakes")
```

captions that explain the labeling on the points and on the y-axis. It will be necessary to explain how distances on the scale relate to changes in area.

### (b) Repeat the plot and associated labeling, now plotting area versus elevation, but specifying log=“y” in order to obtain a logarithmic y-scale.

```{r}
plot(area ~ elevation, pch=16, xlim=c(170,280), ylog=T)
text(area ~ elevation, labels=row.names(Manitoba.lakes), pos=4,ylog=T)
text(area ~ elevation, labels=area, pos=2, ylog=T)
title("Manitoba’s Largest Lakes")
```

#MB.Ch1.7. 

Look up the help page for the R function dotchart(). Use this function to display the areas of the Manitoba lakes (a) on a linear scale, and (b) on a logarithmic scale. Add, in each case, suitable labeling information.

The areas of the Manitoba lakes (a) on a linear scale
```{r}
dotchart(area,labels =row.names(Manitoba.lakes),xlab ="area", ylab = NULL)
```

The areas of the Manitoba lakes (a) on a logarithmic scale

```{r}
dotchart(log2(area), labels =row.names(Manitoba.lakes),xlab ="log2(area)", ylab = NULL)
```


# MB.Ch1.8. Using the sum() function, obtain a lower bound for the area of Manitoba covered by water.
```{r}
sum(area)
```

